The shape of kelp: environmental influences on kelp morphology

### 2018-11-08 chapter_2.Rmd
Edits by AJS.

---
title: "The shape of kelp: environmental influences on kelp morphology"
output:
  pdf_document:
    toc: yes
  html_notebook:
    fig_caption: yes
    toc: yes
  word_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    toc: yes
  bookdown::html_document2:
    fig_caption: yes
    fig_retina: 2
    fig_width: 7
    highlight: espresso
    theme: sandstone
    toc_float: no
fontsize: 12pt
geometry: margin=1in
language: Australian
mainfont: PT Serif
monofont: PT Mono
csl: LaTeX/ecography.csl
papersize: A4
sansfont: PT Sans
tables: yes
bibliography: LaTex/mycollection.bib
---

@@

Kelps are a group of large seaweeds of the order Laminariales (Ochrophyta), which despite their relatively low taxonomic diversity of species in genera [@Bolton2010], nevertheless form the basis of one of the most productive ecosystems globally [@Mann1973]. Kelps generally have a dependence on cool, temperate and arctic seawater temperatures [@Santelices2007; @Bolton2010], and dominate the nearshore biomass within the rocky shallow coasts in both hemispheres [@Steneck2002]. Outside of these latitudes, they are also found in cool, deep water towards the tropics [@Graham2007]: this deeper tropical water, due to its water clarity, also allows for suitable photon flux and a nitrate concentration [@Luning1990; @Dayton1999; @Zimmerman1984], the two other environmental variables that support kelp productivity. Although kelps have a low taxonomic diversity, their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude other turf and subcanopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates [@Mann1973; @Duggins1989; @Steneck2002], each depending on a wide suite of ecological services provided by the kelp habitat [@Gaines1987; @Bologna1993; @Levin1994; @Anderson1997].

@@

Kelps are sessile species and have to adapt to local environmental conditions in order to persist and survive. The resilience of kelps is evident through their distribution, which spans a range of environmental conditions that vary spatially and temporally. Wave exposure and temperature are regarded as important environmental drivers with regards to macroalgae, and play a role in the distribution, abundance, diversity, composition, growth [@Cousens1982] and productivity [@pedersen2012a] of kelps. Wave exposure is recognised as the main cause of kelp mortality, which may manifest itself through dislodgment of individuals or breakage of important morphological structures. These mechanisms usually occur during times of high wave energy or pulse disturbance events, such as storms.Mechanical forces generated by the hydrodynamic environment, in the form of sudden strong ocean currents or storms, between 10- 20 m s-1 with accelerations of 400 m s-2 [@Friedland1995] are the biggest threat to kelp survival. 
Kelps are able to reduce the probability of dislodgment or breakage through morphological adaptation, which may involve several strategies. Kelps can either reduce drag, increase attachment strength or increase flexibility in order to survive high wave energy environments.The strategy that the algal organism employs is dependent on the species. For instance, branched, articulated algae will vary its morphology differently to that of foliose, more flexible macroalgae. For example, @moss1948 investigated the anatomy, chemical composition of *Fucus spiralis* at three sites that varied in wave exposure (sheltered, medium exposure and exposed). The authors found that individuals in exposed sites showed less branching of thalli as well as variation physiological components, such as organic nitrogen, mannitol, laminarian and alginic acid concentrations. The authors also noted a ‘crumpling effect’ displayed by individuals from exposed sites and inferred that this strategy may reduce overall drag.  Other studies show that macroalgae in wave exposed environments have morphologies that reduce overall drag, increase strength of attachment or increase flexibility. For example, a study by @FowlerWalker2006 tested for differences in morphology of *Ecklonia radiata* between wave-sheltered and wave-exposed sites and through a combination of *in situ* sampling and transplantation of juvenile plants. The results showed that morphology differed between wave-sheltered and wave-exposed sites (thin thallus at sheltered sites and a narrow, thick thallus with a thick stipe at exposed sites), and was consistent with previous studies. Juveniles transplanted into wave exposed sites under went rapid morphological adaptation, whilst the opposite was true for wave-sheltered sites which showed slower morphological adaptation. Dislodgement through increased wave exposure is not the only hydrological parameter playing a role in kelp mortality. A study by @DeBettignies2013 tested if there is a linked between orbital water velocity representative of wave-swept reefs and the morphology and biomass of *Ecklonia radiata*. The results showed that morphological adaptation through drag reduction and strength increase strategies do not reduce drag as efficiently as reduction in overall biomass through reduction in thallus surface area. The authors concluded that at peak orbitat water velocities size (biomass) and not morphological adaptation was the main driver of hydrological performance in wave-swept environments.

There is also evidence that morphological adaptation is driven by currents, and in fact may be driving hydrological performance of macrolgae. @Duggins2003 examined the direct and indirect flow effects on population dynamics, morphology and biomechanics of several understorey macroalgae species. These species included *Costaria costata*, *Agarum fimbriatum*, and *Laminaria complanata* and * Nereocystis luetkeana*. The results showed that in wave impacted sites (wave exposed) had higher rates of mortality, and no significance was found between survival of individuals and tidal or current velocity. The authors concluded that although tidal and current velocity did not play a significant role in determining kelp survival, it did play a role in morphological adaptation. The results from this study suggest that high current and tidal stresses are the main driver of kelp morphological adaptation. This in turn make those individuals more resilient to dislodgement to wave exposure. Wave exposure is stochastic in nature compared to tidal and ocean currents which are more regular in their frequency and magnitude. Therefore the regular forces of tidal and ocean currents may make kelp individuals more resilient to mechanical dislodgement over time.

@@

```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="A map indicating the sites where morphometric measurements were collected for _Ecklonia maxima_ and _Laminaria pallida_.", fig.height=8, fig.width=12}

@@

Characterised by a large distal swollen bulb filled with gas, and smooth fronds, this species grows to approximately 10 meters [@Bolton1987]. There was, however, a 17-meter specimen collected in 2015 off Cape Point (Smit, unpubl. data).This species of kelp not only dominate the biomass of the South African nearshore, but plays an important ecological role [@Bustamante1996]. The estimated productivity of _Ecklonia maxima_ within South Africa varies between 350 and 1500g Cm-2yr-1 [@Mann1982]. Across the majority of the coastline, _Laminaria pallida_ remains a subsurface kelp, dominating the kelp biomass at depths greater than 10 meters [@Field1980b; @Bolton1987; @Molloy1996]. This species is distributed from Danger Point, east of the Cape Peninsula, to Rocky Point in northern Namibia, and reaches depths of greater than 20 meters [@Field1980b; @Molloy1996; @Molloy1996; @Stegenga1997]. Towards the north along the west coast, from around Hondeklipbaai, _Laminaria pallida_ replaces _Ecklonia maxima_ as the dominant kelp species [@Velimirov1977; @Stegenga1997] and it also occupies increasingly shallower subtidal regions. The northern populations also exhibit an increase in stipe hollowness, compared to the solid stipe morphs in the species’ southern distributions [@Molloy1996]. This variation in morphology was thought to represent two distinct species, with the northern populations formerly described as _Laminaria schinzii_ Foslie [@Molloy1996]. Genetic work has subsequently shown that the two morphs are in fact the same species [@Rothman2017]. In southern African waters, the primary production of _Laminaria pallida_ is between 120 and 1900g C m2yr1, similar to that of _Ecklonia maxima_ [@Mann1982]. Primary production is not the only pathway.

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Coastal wave energy exposure around the Cape Peninsula for the (a) total 17-year nearshore wave record, (b) only winter months and (c) only summer months."}

@@

grid.arrange(img_we_total, img_we_winter,img_we_summer, ncol = 2,nrow = 2)

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="SWAN numerical output for an example typical offshore wave conditions refracting around the Cape Peninsula."}

knitr::include_graphics('figures/regional_waves.png')

@@

The temperature parameters around the Western Cape coast vary among sites and seasons. During February and August, the west side of Cape point experiences decreased temperatures (Fig. 5).  This region also experiences lower mean and maximum temperatures when compared to False Bay, a region known as east of Cape Point. The range of temperatures within False Bay are larger in winter (August) compared to summer (February) (Fig. 5). The sites north of Kommetjie on the west side display larger temperatures ranges than sites south of Kommetjie on the west side.

@@

```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Boxplots representing the different _Laminaria pallida_ morphometrics measured around the Western Cape coastline, with the X-axis depicting the specific morphology measured, with units provided. Boxplots represent the minimum, 25^th^ percentile, median and 75^th^ percentile of the morphometrics measured. Interquartile range can be deduced as the different between the 75^th^ and 25^th^ percentiles, and dots represent outliers in the data. Sites are ordered sequentially on the Y-axis by location along the coast. The top site is Sea Point and is located at the north western boundary, and Betty’s Bay as the bottom site is located at the south eastern boundary, from our sample region.", fig.height=8, fig.width=12, fig.pos="H"}

@@

```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Boxplots representing the different _Ecklonia maxima_ morphometrics measured around the Western Cape coastline, with the X-axis depicting the specific morphology measured, with units provided. Boxplots represent the minimum, 25^th^ percentile, median and 75^th^ percentile of the morphometrics measured. Interquartile range can be deduced as the different between the 75^th^ and 25^th^ percentiles, and dots represent outliers in the data. Sites are ordered sequentially on the Y-axis by location along the coast. The top site is St. Helena Bay and is located at the north western boundary, and Betty’s Bay as the bottom site is located at the south eastern boundary, from our sample region.", fig.height=8, fig.width=12, fig.pos="H"}

@@

```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Boxplots representing the different _Ecklonia maxima_ morphometrics between deep and shallow water sites measured around the Western Cape coastline, with the X-axis depicting the specific morphology measured, with units provided. Boxplots represent the minimum, 25^th^ percentile, median and 75^th^ percentile of the morphometrics measured. Interquartile range can be deduced as the different between the 75^th^ and 25^th^ percentiles, and dots represent outliers in the data. Sites are ordered sequentially on the Y-axis by location along the coast. The top site is St. Helena Bay and is located at the north western boundary, and Betty’s Bay as the bottom site is located at the south eastern boundary, from our sample region.", fig.height=8, fig.width=12, fig.pos="H"}

@@

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide", fig.cap="A correlation graph depicting the relationships that various temperature and wave parameters share, at an annual timescale. The left Y-axis as well as bottom X-axis represent the measurements for the parameters, while the top X-axis and right Y-axis depict the abiotic parameter name. The top right section of the graph represents correlation coefficients between abiotic parameters. The bottom left section provides individual linear regressions between each abiotic parameter, with a fitted line. Each point of each linear graph represents a collection site. The diagonal density plots represent the spread of the data for each abiotic relationship.", fig.height=8, fig.width=12, fig.pos="H"}
# # ggpairs to plot correlation coefficients
# # Annual
ggpairs(data = ann_abio_cor, columns = 2:14, 
        upper = list(continuous = wrap("cor", size = 3)), 
        lower = list(continuous = "smooth"), 
        diag = list(continuous = "densityDiag")) + 
  theme(text = element_text(size = 9)) + 
  ggtitle("Annual")

@@

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide", fig.cap="A correlation graph depicting the relationships that various temperature and wave parameters share, during February. The left Y-axis as well as bottom X-axis represent the measurements for the parameters, while the top X-axis and right Y-axis depict the abiotic parameter name. The top right section of the graph represents correlation coefficients between abiotic parameters. The bottom left section provides individual linear regressions between each abiotic parameter, with a fitted line. Each point of each linear graph represents a collection site. The diagonal density plots represent the spread of the data for each abiotic relationship.", fig.height=8, fig.width=12, fig.pos="H"}

@@

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide", fig.cap="An RDA, represented by the first two RDA axes, depict the influence that seasonal temperature parameters have on the morphology of _Ecklonia maxima_ sporophytes. Environmental (explanatory) variables, in this case seasonal temperature parameters, are represented by blue vectors extending from the origin. Response variables, in this case _Ecklonia maxima_ morphologies, are representing by red points ordinated across the plane. Sites are represented in black. The top X-axis and right Y-axis represent the explanatory variable axes, while bottom X-axis and left Y-axis represent the response variables axes.", fig.height=8, fig.width=12, fig.pos="H"}
# Plot scaling = 2
plot(eck_temp_season_RDA, scaling = 2)

@@

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide", fig.cap="An RDA, represented by the first two RDA axes, depict the influence that annual wave parameters have on the morphology of _Laminaria pallida_ sporophytes. Environmental (explanatory) variables, in this case annual wave parameters, are represented by blue vectors extending from the origin. Response variables, in this case _Laminaria pallida_ morphologies, are representing by red points ordinated across the plane. Sites are represented in black. The top X-axis and right Y-axis represent the explanatory variable axes, while bottom X-axis and left Y-axis represent the response variables axes.", fig.height=8, fig.width=12, fig.pos="H"}
# Plot scaling = 2
plot(lam_wave_RDA, scaling  = 2)

@@

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide", fig.cap="An RDA, represented by the first two RDA axes, depict the influence that seasonal temperature parameters have on the morphology of _Laminaria pallida_ sporophytes. Environmental (explanatory) variables, in this case seasonal temperature parameters, are represented by blue vectors extending from the origin. Response variables, in this case _Laminaria pallida_ morphologies, are representing by red points ordinated across the plane. Sites are represented in black. The top X-axis and right Y-axis represent the explanatory variable axes, while bottom X-axis and left Y-axis represent the response variables axes.", fig.height=8, fig.width=12, fig.pos="H"}

@@

Kelps, in particular brown seaweeds, are robust and resilient organisms that have been shown to adapt their morphology to suit local environmental conditions. Wave exposure has been shown to be the main driver of morphological variation in brown seaweeds as a strategy to reduce drag and ultimately dislodgement. Although wave exposure is regarded as an important abiotic variable that drives kelp morphology it does not act independently. For instance, a study by @Wernberg2005 investigated the effect of wave exposure on the morphology of _Ecklonia radiata_ of six locations along 1100 km of the southwest Australian coastline. The authors concluded that wave exposure was the main driver of kelp morphological adaptation in the study, however the authors also noted the variation in morphology between sites was not consistent and concluded that wave parameters may play different roles in determining kelp morphology. The results from this current study confirm the idea that kelp morphology is driven by specific wave parameters, particularly significant wave height, wave period and wave direction.

>>>>

---
title: "The shape of kelp: environmental influences on kelp morphology"
author:
- affiliation: University of the Western Cape
  name: Ross Coppin
date: "8 November 2018"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 6.5
    fig_width: 6.5
    highlight: default
    keep_tex: yes
    latex_engine: xelatex
    template: LaTeX/template-article_ajs.tex
  html_notebook:
    fig_caption: yes
    toc: yes
  word_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    toc: yes
  bookdown::html_document2:
    fig_caption: yes
    fig_retina: 2
    fig_width: 7
    highlight: espresso
    theme: sandstone
    toc_float: no
tables: yes
fontsize: 10pt
papersize: A4
geometry: margin=1in
language: Australian
csl: LaTeX/ecography.csl
bibliography: LaTeX/mycollection.bib
---

@@

Kelps are a group of large seaweeds of the order Laminariales (Ochrophyta), which despite their relatively low taxonomic diversity of species in genera [@Bolton2010], nevertheless form the basis of one of the most productive ecosystems globally [@Mann1973]. Kelps generally have a dependence on cool, temperate and arctic seawater temperatures [@Santelices2007; @Bolton2010], and dominate the nearshore biomass within the rocky shallow coasts in both hemispheres [@Steneck2002]. Outside of these latitudes, they are also found in cool, deep water towards the tropics [@Graham2007]: this deeper tropical water, due to its water clarity, also allows for suitable photon flux and a nitrate concentration [@Luning1990; @Dayton1999; @Zimmerman1984], the two other environmental variables that support kelp productivity. Although kelps have a low taxonomic diversity, their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude other turf and subcanopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates [@Mann1973; @Duggins1989; @Steneck2002], each depending on a wide suite of ecological services provided by the kelp and the habitat structure they provide [@Gaines1987; @Bologna1993; @Levin1994; @Anderson1997].

@@

Kelps are sessile species and have to adapt to local environmental conditions in order to grow, repruce and persist. The resilience of kelps is evident through their distribution, which spans a range of environmental conditions that vary spatially and temporally. Wave exposure and temperature are regarded as important environmental drivers with regards to macroalgae, and play a role in the distribution, abundance, diversity, composition, growth [@Cousens1982] and productivity [@pedersen2012a] of kelps. Wave exposure is recognised as the main cause of kelp mortality, which may manifest itself through dislodgment of individuals or breakage of important morphological structures. These mechanisms usually occur during times of high wave energy or pulse disturbance events, such as storms. Mechanical forces generated by the hydrodynamic environment, in the form of sudden strong ocean currents or storms, between 10-20 m.s^-1^ with accelerations of 400 m.s^-1^ [@Friedland1995] are the biggest threat to kelp survival. 
Kelps are able to reduce the probability of dislodgment or breakage through morphological adaptation, which may involve several strategies. Kelps can either reduce drag, increase attachment strength, or increase flexibility in order to survive high wave energy environments. The strategy that the algal organism employs is dependent on the species. For instance, branched, articulated algae will vary its morphology differently to that of foliose, more flexible macroalgae. @moss1948 investigated the anatomy and chemical composition of *Fucus spiralis* at three sites that varied in wave exposure (sheltered, medium exposure and exposed). The author found that individuals in exposed sites showed less branching of thalli as well as variation physiological components, such as organic nitrogen, mannitol, laminarian and alginic acid concentrations. The authors also noted a ‘crumpling effect’ displayed by individuals from exposed sites and inferred that this strategy may reduce overall drag.  Other studies show that macroalgae in wave exposed environments have morphologies that reduce overall drag, increase strength of attachment or increase flexibility. For example, a study by @FowlerWalker2006 looked for differences in morphology of *Ecklonia radiata* between wave-sheltered and wave-exposed sites, using a combination of *in situ* sampling and transplantation of juvenile plants. The results showed that morphology differed between wave-sheltered and wave-exposed sites, with thin thalli developing at sheltered sites and a narrower, thicker thallus with a thick stipe forming at exposed sites. This finding was consistent with previous studies {AJS: name the studies and the species involved here}. Juveniles transplanted into wave exposed sites underwent rapid morphological adaptation, whilst the opposite was true for wave-sheltered sites, which showed slower morphological adjustment {refs.}. Dislodgement through increased wave exposure is not the only hydrological variable playing a role in kelp mortality. A study by @DeBettignies2013 tested if there is a link between orbital water velocity representative of wave-swept reefs and the morphology and biomass of *Ecklonia radiata*. The results showed that morphological adaptation through drag reduction and strength increase does not reduce drag as efficiently as reduction in overall biomass through reduction in thallus surface area. The authors concluded that at peak orbital water velocities, size (as biomass) and not morphological adaptation was the main determinant of hydrological performance in wave-swept environments.

There is also evidence that morphological adaptation is driven by currents, and in fact may be driving hydrological performance of macrolgae. @Duggins2003 examined the direct and indirect flow effects on population dynamics, morphology and biomechanics of several understorey macroalgae species. These species included *Costaria costata*, *Agarum fimbriatum*, *Laminaria complanata* and *Nereocystis luetkeana*. The results showed that wave impacted sites (wave exposed) exhibited higher rates of mortality, and no significance was found between survival of individuals and tidal or current velocity. The authors concluded that although tidal and current velocity did not play a significant role in determining kelp survival, it did play a role in morphological adaptation. The results from this study suggest that high current and tidal stresses are the main driver of kelp morphological adaptation. This in turn make those individuals more resilient to dislodgement to wave exposure. Wave exposure is stochastic in nature compared to tidal and ocean currents which are more regular in their frequency and magnitude. Therefore the regular forces of tidal and ocean currents may make kelp individuals more resilient to mechanical dislodgement over time.

@@

```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="A map indicating the sites where morphometric measurements were collected for _Ecklonia maxima_ and _Laminaria pallida_."}

@@

Characterised by a large distal swollen bulb filled with gas, and smooth fronds, this species grows to approximately 10 meters [@Bolton1987]. There was, however, a 17 m specimen collected in 2015 off Cape Point (Smit, unpubl. data). This species of kelp not only dominate the biomass of the South African nearshore, but plays an important ecological role [@Bustamante1996]. The estimated productivity of _Ecklonia maxima_ within South Africa varies between 350 and 1500 gC.m^-2^yr^-1^ [@Mann1982]. Across the majority of the coastline, _Laminaria pallida_ remains a subsurface kelp, dominating the kelp biomass at depths greater than 10 m [@Field1980b; @Bolton1987; @Molloy1996]. This species is distributed from Danger Point, east of the Cape Peninsula, to Rocky Point in northern Namibia, and reaches depths of greater than 20 m [@Field1980b; @Molloy1996; @Molloy1996; @Stegenga1997]. Towards the north along the west coast, from around Hondeklipbaai, _Laminaria pallida_ replaces _Ecklonia maxima_ as the dominant kelp species [@Velimirov1977; @Stegenga1997] and it also occupies increasingly shallower subtidal regions. The northern populations also exhibit an increase in stipe hollowness, compared to the solid stipe morphs in the species’ southern distributions [@Molloy1996]. This variation in morphology was thought to represent two distinct species, with the northern populations formerly described as _Laminaria schinzii_ Foslie [@Molloy1996]. Genetic work has subsequently shown that the two morphs are in fact the same species [@Rothman2017]. In southern African waters, the primary production of _Laminaria pallida_ is between 120 and 1900g C m2yr1, similar to that of _Ecklonia maxima_ [@Mann1982]. Primary production is not the only pathway.

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Coastal wave energy exposure around the Cape Peninsula for the (a) total 17-year nearshore wave record, (b) only winter months and (c) only summer months."}

@@

grid.arrange(img_we_total, img_we_winter, img_we_summer, ncol = 2, nrow = 2)

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "SWAN numerical output for an example typical offshore wave conditions refracting around the Cape Peninsula."}

thing <- rasterGrob(as.raster(readPNG("figures/regional_waves.png")), interpolate = FALSE)
grid.arrange(thing, ncol = 1)
# knitr::include_graphics('figures/regional_waves.png')

@@

The temperature parameters around the Western Cape coast vary among sites and seasons. During February and August, the west side of Cape Point experiences decreased temperatures (Fig. 5). This region also experiences lower mean and maximum temperatures when compared to False Bay, a region known as east of Cape Point. The range of temperatures within False Bay are larger in winter (August) compared to summer (February) (Fig. 5). The sites north of Kommetjie on the west side display larger temperatures ranges than sites south of Kommetjie on the west side.

@@

```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Boxplots representing the different _Laminaria pallida_ morphometrics measured around the Western Cape coastline, with the X-axis depicting the specific morphology measured, with units provided. Boxplots represent the minimum, 25^th^ percentile, median and 75^th^ percentile of the morphometrics measured. Interquartile range can be deduced as the different between the 75^th^ and 25^th^ percentiles, and dots represent outliers in the data. Sites are ordered sequentially on the Y-axis by location along the coast. The top site is Sea Point and is located at the north western boundary, and Betty’s Bay as the bottom site is located at the south eastern boundary, from our sample region.", fig.pos="H"}

@@

```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Boxplots representing the different _Ecklonia maxima_ morphometrics measured around the Western Cape coastline, with the X-axis depicting the specific morphology measured, with units provided. Boxplots represent the minimum, 25^th^ percentile, median and 75^th^ percentile of the morphometrics measured. Interquartile range can be deduced as the different between the 75^th^ and 25^th^ percentiles, and dots represent outliers in the data. Sites are ordered sequentially on the Y-axis by location along the coast. The top site is St. Helena Bay and is located at the north western boundary, and Betty’s Bay as the bottom site is located at the south eastern boundary, from our sample region.", fig.pos="H"}

@@

```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Boxplots representing the different _Ecklonia maxima_ morphometrics between deep and shallow water sites measured around the Western Cape coastline, with the X-axis depicting the specific morphology measured, with units provided. Boxplots represent the minimum, 25^th^ percentile, median and 75^th^ percentile of the morphometrics measured. Interquartile range can be deduced as the different between the 75^th^ and 25^th^ percentiles, and dots represent outliers in the data. Sites are ordered sequentially on the Y-axis by location along the coast. The top site is St. Helena Bay and is located at the north western boundary, and Betty’s Bay as the bottom site is located at the south eastern boundary, from our sample region.", fig.pos="H"}

@@

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide", fig.cap="A correlation graph depicting the relationships that various temperature and wave parameters share, at an annual timescale. The left Y-axis as well as bottom X-axis represent the measurements for the parameters, while the top X-axis and right Y-axis depict the abiotic parameter name. The top right section of the graph represents correlation coefficients between abiotic parameters. The bottom left section provides individual linear regressions between each abiotic parameter, with a fitted line. Each point of each linear graph represents a collection site. The diagonal density plots represent the spread of the data for each abiotic relationship.", fig.pos="H"}
# # ggpairs to plot correlation coefficients
# # Annual
ggpairs(data = ann_abio_cor, columns = 2:14, 
        upper = list(continuous = wrap("cor", size = 3)), 
        lower = list(continuous = "smooth"), 
        diag = list(continuous = "densityDiag")) + 
  theme(text = element_text(size = 9)) + 
  ggtitle("Annual")

@@

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide", fig.cap="A correlation graph depicting the relationships that various temperature and wave parameters share, during February. The left Y-axis as well as bottom X-axis represent the measurements for the parameters, while the top X-axis and right Y-axis depict the abiotic parameter name. The top right section of the graph represents correlation coefficients between abiotic parameters. The bottom left section provides individual linear regressions between each abiotic parameter, with a fitted line. Each point of each linear graph represents a collection site. The diagonal density plots represent the spread of the data for each abiotic relationship.", fig.pos="H"}

@@

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide", fig.cap="An RDA, represented by the first two RDA axes, depict the influence that seasonal temperature parameters have on the morphology of _Ecklonia maxima_ sporophytes. Environmental (explanatory) variables, in this case seasonal temperature parameters, are represented by blue vectors extending from the origin. Response variables, in this case _Ecklonia maxima_ morphologies, are representing by red points ordinated across the plane. Sites are represented in black. The top X-axis and right Y-axis represent the explanatory variable axes, while bottom X-axis and left Y-axis represent the response variables axes.", fig.pos="H"}
# Plot scaling = 2
plot(eck_temp_season_RDA, scaling = 2)

@@

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide", fig.cap="An RDA, represented by the first two RDA axes, depict the influence that annual wave parameters have on the morphology of _Laminaria pallida_ sporophytes. Environmental (explanatory) variables, in this case annual wave parameters, are represented by blue vectors extending from the origin. Response variables, in this case _Laminaria pallida_ morphologies, are representing by red points ordinated across the plane. Sites are represented in black. The top X-axis and right Y-axis represent the explanatory variable axes, while bottom X-axis and left Y-axis represent the response variables axes.", fig.pos="H"}
# Plot scaling = 2
plot(lam_wave_RDA, scaling  = 2)

@@

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide", fig.cap="An RDA, represented by the first two RDA axes, depict the influence that seasonal temperature parameters have on the morphology of _Laminaria pallida_ sporophytes. Environmental (explanatory) variables, in this case seasonal temperature parameters, are represented by blue vectors extending from the origin. Response variables, in this case _Laminaria pallida_ morphologies, are representing by red points ordinated across the plane. Sites are represented in black. The top X-axis and right Y-axis represent the explanatory variable axes, while bottom X-axis and left Y-axis represent the response variables axes.", fig.pos="H"}

@@

Kelps, in particular brown seaweeds, are robust and resilient organisms that have been shown to adapt their morphology to suit local environmental conditions (refs.). Wave exposure has been shown to be the main driver of morphological variation in brown seaweeds as a strategy to reduce drag and ultimately dislodgement (refs.). Although wave exposure is regarded as an important abiotic variable that drives kelp morphology, it does not act independently {AJS: of what?}. For instance, a study by @Wernberg2005 investigated the effect of wave exposure on the morphology of _Ecklonia radiata_ of six locations along 1100 km of the southwest Australian coastline. The authors concluded that wave exposure was the main driver of kelp morphological adaptation in the study; however, they also noted the variation in morphology between sites was not consistent and concluded that wave parameters may play different roles in determining kelp morphology {AJS: What does this mean?}. The results from this current study confirm the idea that kelp morphology is driven by specific wave parameters, particularly significant wave height, wave period and wave direction {AJS: what about?}.
